---
title: "Final_Project"
author: "Andrew Ross"
date: "`r Sys.Date()`"
output: html_document
---

The purpose of this script is to analyze data to predict quaternions with different covariates using
Multivariate analysis 

Import Libraries 


```{r message=FALSE, warning=FALSE}
library(dplyr) # For data manipulation
library(ggplot2) # for nice visuals
library(ggpubr) # For nicer normality plots
library(car) #For Transformations
library(MVN)
library(tidyr)
```

Import Dataset


```{r}
data1 = read.csv("D:\\Stat 9370\\IMU_Data_1.csv")
data2 = read.csv("D:\\Stat 9370\\IMU_Data_2.csv")
data3 = read.csv("D:\\Stat 9370\\IMU_Data_3.csv")
data4 = read.csv("D:\\Stat 9370\\IMU_Data_4.csv")
test= read.csv("D:\\Stat 9370\\IMU_Data_5.csv") #For testing the model
```

```{r}
data = rbind(data1,data2,data3,data4) #combine the data into one

data = data |> select(-Euler_x, -Euler_y, -Euler_z) # Removing Euler because professor said so

#str(data)
```

16 variables with 120000 observations that are all numeric.

# EDA

# Univariate -------------------------------------------------------------------------------------------


```{r}
for(i in colnames(data)){
  print(ggplot(data = data, mapping = aes_string(i)) + geom_histogram()+ labs(title = i))
} #This loop plots histograms for each variable
```

```{r}

annotate_figure(
  ggarrange(
    
    ggplot(data = data, mapping = aes(x=Quat_0))+geom_histogram() + labs(title = '0') + labs(x = '', y = '') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Quat_1))+geom_histogram() + labs(title = '1')+labs(x = '',y = '')+
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Quat_2))+geom_histogram() + labs(title = '2')+labs(x = '',y = '')+
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Quat_3))+geom_histogram() + labs(title = '3')+labs(x = '',y = '')+
      theme(plot.title = element_text(hjust = 0.5))
  
  ), left = text_grob('Count', rot = 90), top = text_grob('Quaternion', face = 'bold', size = 14)
)
```


```{r, fig.width=9}
annotate_figure(
  ggarrange(
  
    ggplot(data = data, mapping = aes(x=Gyro_x))+geom_histogram() + labs(title = 'x', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Gyro_y))+geom_histogram() + labs(title = 'y', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Gyro_z))+geom_histogram() +labs(title = 'z', x = '', y ='')+
      theme(plot.title = element_text(hjust = 0.5)),
    
    ncol = 3

  ) , left = text_grob('Count', size = 10, rot = 90), top = text_grob('Gyro', size = 14, face = 'bold')
)

```


```{r}
annotate_figure(
  ggarrange(
  
    ggplot(data = data, mapping = aes(x=Acc_x))+geom_histogram() + labs(title = 'x', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Acc_y))+geom_histogram() + labs(title = 'y', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Acc_z))+geom_histogram() +labs(title = 'z', x = '', y ='')+
      theme(plot.title = element_text(hjust = 0.5)),
    
    ncol = 3

  ) , left = text_grob('Count', size = 10, rot = 90), top = text_grob('Accelerometer', size = 14, 
                                                                      face = 'bold')
)
```


```{r fig.width=7}
annotate_figure(
  ggarrange(
  
    ggplot(data = data, mapping = aes(x=Mag_x))+geom_histogram() + labs(title = 'x', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Mag_y))+geom_histogram() + labs(title = 'y', x = '', y ='') +
      theme(plot.title = element_text(hjust = 0.5)),
    
    ggplot(data = data, mapping = aes(x=Mag_z))+geom_histogram() +labs(title = 'z', x = '', y ='')+
      theme(plot.title = element_text(hjust = 0.5)),
    
    ncol = 3

  ) , left = text_grob('Count', size = 10, rot = 90), top = text_grob('Magnetometer', size = 14, face = 'bold')
)
```




 Each Gyro, each acc are visually normally distributed


# transformations

Box cox won't work for this because I have negative values in my variables so we try yeo johnson


```{r}
b = boxCox(lm(Mag_x~1, data = data), family="yjPower", plotit = TRUE)
b$x[which.max(b$y)]

b = boxCox(lm(Mag_y~1, data = data), family="yjPower", plotit = TRUE)
b$x[which.max(b$y)]

b = boxCox(lm(Mag_z~1, data = data), family="yjPower", plotit = TRUE)
b$x[which.max(b$y)]

```

```{r}
hist(data$Mag_x)
hist(yjPower(data$Mag_x, lambda = 1.030303))

hist(data$Mag_y)
hist(yjPower(data$Mag_y, lambda = 1.030303))

hist(data$Mag_z)
hist(yjPower(data$Mag_z, lambda = 0.989899))


yjmx = yjPower(data$Mag_x, lambda = 1.030303)
yjmy = yjPower(data$Mag_y, lambda = 1.030303)
yjmz = yjPower(data$Mag_z, lambda = 0.989899)

```


# Might need to think of different transformation with mag

# Now we proceed with quat
```{r}
boxCox(lm(Quat_0~1, data = data), family="yjPower", plotit = TRUE)
boxCox(lm(Quat_1~1, data = data), family="yjPower", plotit = TRUE)
boxCox(lm(Quat_2~1, data = data), family="yjPower", plotit = TRUE)

```


```{r}
hist(yjPower(data$Quat_0, lambda = 1))
qqPlot(yjPower(data$Quat_0, lambda = 1))
ggqqplot(yjPower(data$Quat_0, lambda = 1))



hist(yjPower(data$Quat_1, lambda = 0.9))
qqPlot(yjPower(data$Quat_1, lambda = 0.9))

hist(yjPower(data$Quat_0, lambda = 1))
qqPlot(yjPower(data$Quat_0, lambda = 1))


```

# Bivariate


```{r}
cor(data) # correlation matrix

GGally::ggcorr(data, label = T, hjust = 0.5, label_size = 3, size = 2.5)
```

welp... no values are correlated with each other.  

```{r}
pairs(data)
```


# More Stuff 


# Modeling

```{r}

mod = lm(cbind(Quat_0, Quat_1, Quat_2, Quat_3)~., data = data)

summary(mod)

hist(mod$residuals, main = 'Density of the Residuals', xlab = 'Residuals')

```


```{r}
Anova(mod)

Anova(yjmod)
```



```{r}
test = test |> select(-Euler_x, -Euler_y, -Euler_z)

obs = test |> select(Quat_0, Quat_1, Quat_2, Quat_3)

obs = obs |> pivot_longer(cols = c(Quat_0, Quat_1, Quat_2, Quat_3), names_to = 'Quat', values_to = 'obs')

pred_mod = predict(mod, newdata = test)

pred_mod = as.data.frame(pred_mod)

pred_mod= pred_mod |> pivot_longer(cols = c(Quat_0, Quat_1, Quat_2, Quat_3), names_to = 'Quat', values_to = 'obs')


mse = mean(abs(obs$obs-pred_mod$obs))

```

















